---
title: "Building micRogRad"
execute: 
  error: true
format: 
  html:
    number-sections: true
    toc: true
    toc-location: left
    toc-depth: 4
    code-tools: true
    code-fold: false
    code-link: true
editor: visual
---

## Building micrograd in R

In this notebook, we will attempt to build an R version of [Andrej Karpathy's micrograd](https://github.com/karpathy/micrograd) in R. `Micrograd` is a tiny scalar-valued autograd engine and a neural net library on top of it with PyTorch-like API.

::: callout-warning
Some fair warning 🤗

This is me trying to learn R's OOP system R7: <https://github.com/RConsortium/OOP-WG>. Might be hacky here and there.

Here's my learning process for R7: <https://rpubs.com/eR_ic/r7>
:::

## Initial class definition

```{r}
# Load R7 library
library(R7)


```

To begin with, we create a class `Value` that takes a scalar, keeps track of it and prints it out

```{r}
# Define a class value that takes a numeric input
Value <- new_class(name = "Value",
                   # Specify properties/data that the object will posess
                   properties = list(data = class_numeric))

# Register a cat method on S3 "print" generic
method(generic = print, Value) <- function(x, ...){
  cat("Value(data=", x@data,")", sep = "")
}


# Create a class instance
a <- Value(2.0)
a
```

::: callout-note
R7 uses **functional OOP** where **methods belong to** **generic** functions. We have registered a **method** on the S3 generic `print` that will implement the `cat` function on our R7 class `Value`
:::

```{r}
a = Value(2.0)
b = Value(-3.0)
b^a
```

Ooops. Not so fast. Let's add some mathematical methods too.

## Add mathematical operators

```{r}
# Define class Value
# Value <- new_class("Value",
#                    properties = list(data = class_numeric))


# Ops.Value = function(e1, e2) get(.Generic)(e1@data, e2@data)
S4_register(Value)
method(Ops, list(Value, NULL)) <- function(e1, e2){
  getGenerics(e1@data, e2@data)}
```
