---
title: "Building micRogRad"
execute: 
  error: true
format: 
  html:
    number-sections: true
    toc: true
    toc-location: left
    toc-depth: 4
    code-tools: true
    code-fold: false
    code-link: true
editor: visual
---

## Building micrograd in R

In this notebook, we will attempt to build an R version of [Andrej Karpathy's micrograd](https://github.com/karpathy/micrograd). `Micrograd` is a tiny scalar-valued autograd engine and a neural net library on top of it with PyTorch-like API.

::: callout-warning
Some fair warning 🤗

This is me trying to learn R's OOP system R7: <https://github.com/RConsortium/OOP-WG>. Might be hacky here and there.

Here's my learning process for R7: <https://rpubs.com/eR_ic/r7>
:::

## Initial class definition

```{r}
# Load R7 library
library(R7)
library(magrittr)


```

To begin with, we create a class `Value` that takes a scalar, keeps track of it and prints it out

```{r}
# Define a class value that takes a numeric input
Value <- new_class(name = "Value",
                   # Specify properties/data that the object will possess
                   properties = list(data = class_numeric))

# Register a cat method for Value on S3 "print" generic
method(generic = print, Value) <- function(x, ...){
  cat("Value(data=", x@data,")", sep = "")
}


# Create a class instance
a <- Value(2.0)
a
```

::: callout-note
R7 uses **functional OOP** where **methods belong to** **generic** functions. We have registered a **method** on the S3 generic `print` that will implement the `cat` function on our R7 class `Value`
:::

```{r}
a = Value(2.0)
b = Value(-3.0)
c=a+b
```

Ooops. Not so fast. The object does not know how to handle an addition operation yet. Let's add some mathematical methods too.

## Add mathematical operators

```{r}
# S4_register(Value)
# method(Ops, list(Value, NULL)) <- function(e1, e2){
#   value = get(.Generic)(e1@data, e2@data)
#   # Convert to Value object
#   return(Value(value))
# }
```

```{r}
method(`+`, list(Value, NULL)) <- function(x, y){
  value = a@data + b@data
  # Convert to Value object
  return(Value(value))
}
```

Now `a + b` should work:

```{r}
a+b
```

We can register other mathematical operations(`-`, `*`, etc) individually like before or we can register them all at once on group generic methods such as [`Ops`](https://stat.ethz.ch/R-manual/R-devel/library/base/html/groupGeneric.html) which contain all the mathematical operations:

```{r}
# Register methods for R7 class on group generic
# Ops.Value <- function(e1, e2){
#   value = get(.Generic)(e1@data, e2@data)
#   return(Value(value))
# }


Ops.Value <- function(e1, e2){
  # Mathematical operation
  value = get(.Generic)(e1@data, e2@data)
  # Operation symbol
  op = get(.Generic) %>% deparse() %>% gsub('[.()"\"]', "", .) %>% gsub('\\w', "", .)
  #op = deparse1(op) |> gsub('[.()"\"]', "", .) |> gsub('\\w', "", .)
  return(op)
}

```

::: callout-note
How would be do this using *R7::method\<-()*
:::

```{r}
# Now we can do multiple operations
a <- Value(2.0)
b <- Value(-3.0)
c <- Value(10.0)
d <- a*b + c
d
```

We need to keep track of operations.

Now next step is to expand our class and methods to keep track of what values produced the results and the corresponding mathematical operation.

```{r}
# Add prev property that takes a list of previous values
Value <- new_class(name = "Value",
                   # Specify properties/data that the object will possess
                   properties = list(
                     data = class_numeric,
                     prev = new_property(class_list, default = list()),
                     op = new_property(class_character, default = "")
                     ))



# Register methods for R7 class on group generic
Ops.Value <- function(e1, e2){
  # Mathematical operation
  value = get(.Generic)(e1@data, e2@data)
  # Operation symbol
  op = get(.Generic) %>% 
    deparse() %>% 
    gsub('[.()"\"]', "", .) %>% gsub('\\w', "", .)
  return(Value(value, prev = list(e1, e2), op = op))
}


# Now we can do multiple operations
a <- Value(2.0)
b <- Value(-3.0)
c <- Value(10.0)
d <- a*b + c
d@data
d@prev
d@op
```

```{r}
# Register an R7 class with S4
S4_register(Value)

# Register Ops methods for R7 class
method(Ops, list(Value, NULL)) <- function(e1, e2){
  #value = 
    #getGenerics(e1@data, e2@data)
  method(Ops)(e1@data, e2@data)
  #return(Value(value))
}
a+b
```

```{r}
# Define class Value
# Value <- new_class("Value",
#                    properties = list(data = class_numeric))


#Ops.Value = function(e1, e2) get(.Generic)(e1@data, e2@data)
S4_register(Value)
method(Ops, list(Value, NULL)) <- function(e1, e2){
  getGenerics(e1@data, e2@data)}
```

https://stackoverflow.com/questions/35902360/r-implement-group-generics-ops-to-enable-comparison-of-s3-objects
